From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/ufin.tex
Update

  The type $\dsum*{x:A}{P(x)}$ is the collection of all the fibers and is called the total space of $P$. The first
  projection ${\pi_1 : \dsum*{x:A}{P(x)} \to A}$ from the total space to the base space $A$ has the structure of a
  fibration, that is, there is a lifting operation (\Cref{fig:lift}) which lifts paths in the base space to paths in the
  total space. Given a path $p : x \id_{A} y$ in the base space, and $u : P(x)$ a point in the fiber over $x$, we have:

  \[
    \term{lift}(u,p) : (x , u) \id_{\dsum{x:A}{P(x)}} (y , \transport{P}{p}{u})
  \]

@@

  \begin{proposition}
    The following are equivalent.
    \begin{enumerate}
      \item $f : A \to B$ is an equivalence.
      \item $f$ has a left and right inverse.
      \item $f$ has contractible fibers.
    \end{enumerate}
  \end{proposition}

@@

\begin{propositionrep}
  \label{prop:isFin}
  For any type $X$, $\isFin[X]$ is a proposition.
\end{propositionrep}

@@

Since $\isFin[X]$ is a predicate on the universe $\UU$, we easily get our univalent subuniverse $\UFin$.

@@

\begin{definition}[$\BAut$]
  The predicate $P(X) \defeq \Trunc[-1]{X \id T}$ picks out exactly those types that are merely equal to $T$, and this
  generates the subuniverse
  $
    \BAut[T] \defeq \Sub{T}.
  $
  We write $T_0 \defeq (T, \trunc{\refl_{T}})$ for the image of the inclusion of $T$ in $\BAut[T]$.
\end{definition}

@@

\begin{definition}
  For any $n : \Nat$, we define $\UFin[n] \defeq \BAut[\Fin[n]]$ to be the univalent subuniverse of $n$-element sets.
  Note that, $\UFin$ can be equivalently seen as the collection of all types of finite cardinality, that is,
  $\UFin \eqv \dsum*{n:\Nat}{\UFin[n]}$.
\end{definition}

@@

\begin{propositionrep}
  \leavevmode
  \begin{enumerate}
    \item If $T$ is an $n$-type, $\BAut[T]$ is an $(n+1)$-type.
    \item For any $T : \UU$, $\BAut[T]$ is 0-connected.
    \item For any $T : \UU$, \( \loopspace[\BAut[T],T_{0}] \eqv \Aut[T] \). \label{lem:loop-deloop}
  \end{enumerate}
\end{propositionrep}

@@

\begin{corollary}
  $\UFin[n]$ is a pointed, connected, 1-groupoid for every $n:\Nat$, and \( \loopspace[\UFin[n],F_{n}] \eqv
  \Aut[\Fin[n]] \). $\UFin$ is a 1-groupoid with connected components for every $n:\Nat$.
\end{corollary}

@@

\begin{proposition}
  $\UFin$ has a symmetric rig structure, with the additive and multiplicative symmetric monoidal structures given by
  $(F_0, \sqcup)$ and $(F_1, \times)$, with corresponding natural isomorphisms $\lambda_{X}$, $\rho_{X}$,
  $\alpha_{X,Y,Z}$, and the braiding isomorphism $\mathcal{B}_{X,Y}$ upto 1-paths in $\UFin$. These isomorphisms satisfy
  the Mac Lane coherence conditions for symmetric monoidal categories, that is, the triangle, pentagon, and hexagon
  identities, and the symmetry of the braiding, upto 2-paths in $\UFin$.
\end{proposition}

>>>>

  The type $\dsum*{x:A}{P(x)}$ is the collection of all the fibers and is called the total space of $P$. The first
  projection ${\pi_1 : \dsum*{x:A}{P(x)} \to A}$ from the total space to the base space $A$ has the structure of a
  fibration, that is, there is a lifting operation (\cref{fig:lift}) which lifts paths in the base space to paths in the
  total space. Given a path $p : x \id_{A} y$ in the base space, and $u : P(x)$ a point in the fiber over $x$, we have:
  \[
    \term{lift}(u,p) : (x , u) \id_{\dsum{x:A}{P(x)}} (y , \transport{P}{p}{u})
  \]

@@

  \begin{lemma}
    The following are equivalent.
    \begin{enumerate}
      \item $f : A \to B$ is an equivalence.
      \item $f$ has a left and right inverse.
      \item $f$ has contractible fibers.
    \end{enumerate}
  \end{lemma}

@@

\begin{lemmarep}
  \label{prop:isFin}
  For any type $X$, $\isFin[X]$ is a proposition.
\end{lemmarep}

@@

Since $\isFin$ is a predicate on the universe $\UU$, we easily get our univalent subuniverse $\UFin$.

@@

\begin{definition}[$\BAut$]
  The predicate $P(X) \defeq \Trunc[-1]{X \id T}$ picks out exactly those types that are merely equal to $T$, and this
  generates the subuniverse
  \[
    \BAut[T] \defeq \Sub{T}.
  \]
  We write $T_0 \defeq (T, \trunc{\refl_{T}})$ for the image of the inclusion of $T$ in $\BAut[T]$.
\end{definition}

@@

\begin{definition}[$\UFin[n]$]
  For any $n : \Nat$, we define $\UFin[n] \defeq \BAut[\Fin[n]]$ to be the univalent subuniverse of $n$-element sets.
  Note that, $\UFin$ can be equivalently seen as the collection of all types of finite cardinality, that is,
  $\UFin \eqv \dsum*{n:\Nat}{\UFin[n]}$.
\end{definition}

@@

\begin{lemmarep}
  \leavevmode
  \begin{enumerate}
    \item If $T$ is an $n$-type, $\BAut[T]$ is an $(n+1)$-type.
    \item For any $T : \UU$, $\BAut[T]$ is 0-connected.
    \item For any $T : \UU$, \( \loopspace[\BAut[T],T_{0}] \eqv \Aut[T] \). \label{lem:loop-deloop}
  \end{enumerate}
\end{lemmarep}

@@

\begin{theorem}
  $\UFin[n]$ is a pointed, connected, 1-groupoid for every $n:\Nat$, and \( \loopspace[\UFin[n],F_{n}] \eqv
  \Aut[\Fin[n]] \). $\UFin$ is a 1-groupoid with connected components for every $n:\Nat$.
\end{theorem}

@@

\begin{theorem}
  $\UFin$ has a symmetric rig structure, with the additive and multiplicative symmetric monoidal structures given by
  $(F_0, \sqcup)$ and $(F_1, \times)$, with corresponding natural isomorphisms $\lambda_{X}$, $\rho_{X}$,
  $\alpha_{X,Y,Z}$, and the braiding isomorphism $\mathcal{B}_{X,Y}$ upto 1-paths in $\UFin$. These isomorphisms satisfy
  the Mac Lane coherence conditions for symmetric monoidal categories, that is, the triangle, pentagon, and hexagon
  identities, and the symmetry of the braiding, upto 2-paths in $\UFin$. \todo{Add rig}
\end{theorem}
